---
title: "did group-time ATT adapter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{did group-time ATT adapter}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`dml_from_did()` converts a post-treatment cell from `did::att_gt()` into a
`dml_did` object inheriting from `dml`. The original ATT and IF are retained;
DML estimates the two sensitivity moments on the same cell sample.

## Versions and supported estimates

Compatibility is checked from saved inputs, the required DRDID interfaces, and
numerical reconstruction of the ATT and ordered IF. Versions are recorded as
diagnostics, not used as acceptance gates. This release is tested with did
**2.5.1** and **2.5.1.902**, and DRDID **1.3.0**. Other versions must provide the
same required information and pass the same checks. Missing fields are not
filled with assumed defaults. A saved MP object can be adapted without did
installed; DRDID remains necessary for reconstruction.

```{r install, eval=FALSE}
# Install missing dependencies; did 2.5.1 from CRAN is supported.
install.packages(c("did", "DRDID"))
```

Supported: balanced panel estimation with `est_method = "dr"`,
`compute_inffunc = TRUE`, constant positive weights and inactive trimming.
Both control groups, both computation modes, anticipation and both base-period
settings are supported. A balanced panel retained after did's cleaning is valid.
The `allow_unbalanced_panel` input alone does not determine the estimator path;
the saved `DIDparams$panel` does.

Not supported: actual unbalanced or repeated-cross-section estimation,
`fix_weights = "varying"` (which invokes an RC estimator), other/custom
estimators, pre-treatment/reference cells, clustering above individuals,
active trimming, or `aggte()` objects. The source bootstrap may be enabled,
but sensitivity inference uses analytic joint IFs, not its bootstrap SEs or
simultaneous critical values.

## Cell reconstruction

The adapter reads `fit$DIDparams$data`; the user need not supply historical raw
data. For each requested `(g,t)` it selects the cohort and eligible controls,
then uses the last saved period strictly before `g - anticipation` as the base.
The period grid need not be evenly spaced. Fast mode uses the saved covariate
tensors. Slow mode evaluates the formula on all saved data before selecting
base-period rows, preserving global factor levels and transform bases. Its
formula environment and contrast settings must remain compatible with the fit.
Do not edit the saved source data or tensors.

If the source has `N` units and the cell has `n` units, did stores
`(N / n) * IF_cell` in the cell's column. The adapter matches IF row names to
unit IDs and multiplies by `n / N`. Cell membership comes from the saved data,
not from testing which IF entries are nonzero. DRDID reconstruction must reproduce
both the ATT and the ordered IF before DML runs.

The three IF vectors (ATT, residual variance, imbalance) determine the joint
covariance and sensitivity-bound standard errors. Statistical validity still
requires the estimators' identification, overlap and nuisance conditions.
Reconstruction checks do not establish those assumptions.

## Runnable example

The parametric learners below keep the example small. Choose learners and
folds appropriate to your application. Active DML trimming is rejected; setting
`ps.trim = 0` still requires predictions strictly between zero and one.

```{r setup, include=FALSE}
has_adapter_deps <- requireNamespace("did", quietly = TRUE) &&
  requireNamespace("DRDID", quietly = TRUE)
knitr::opts_chunk$set(eval = has_adapter_deps, collapse = TRUE, comment = "#>")
```

```{r example}
# Tested with did 2.5.1 / 2.5.1.902 and DRDID 1.3.0; acceptance is based on input checks.
library(dml.sensemakr)

set.seed(608)
n <- 360L
units <- data.frame(id = sample(seq_len(n) * 7 + 100),
                    g = rep(c(0, 5, 9), each = n / 3),
                    x = runif(n, -1, 1), z = runif(n, -1, 1), u = rnorm(n))
panel_data <- do.call(rbind, lapply(c(0, 2, 5, 9, 12), function(t) {
  d <- units
  d$time <- t
  d$x <- d$x + 0.04 * t * d$z
  d$y <- d$u + t / 8 + (0.3 + t / 7) * d$x + 0.7 * d$z +
    (d$g > 0 & t >= d$g) * (0.8 + t / 20) + rnorm(n)
  d
}))

source_fit <- did::att_gt(
  yname = "y", tname = "time", idname = "id", gname = "g",
  xformla = ~ x + z, data = panel_data, panel = TRUE,
  est_method = "dr", control_group = "notyettreated",
  bstrap = FALSE, cband = FALSE, compute_inffunc = TRUE
)
args <- list(yreg = "lm", dreg = "glm", d.class = TRUE,
              cf.folds = 3, cf.reps = 3, cf.seed = 17, ps.trim = 0)

# One cell. No raw-data argument is needed.
hybrid <- dml_from_did(source_fit, group = 5, time = 9, dml_args = args)
summary(hybrid)
confidence_bounds(hybrid, cf.y = 0.05, cf.d = 0.05)
sensitivity <- sensemakr(hybrid, cf.y = 0.05, cf.d = 0.05)
benchmark <- dml_benchmark(hybrid, "z")
benchmark_bounds(hybrid, benchmark)

# All post-treatment cells. Inspect status before using individual results.
batch <- dml_from_did_cells(source_fit, dml_args = args)
print(batch)
intervals <- did_cell_apply(batch, confidence_bounds, cf.y = 0.05, cf.d = 0.05)
print(intervals)
intervals$values[["g=5,t=9"]]

# Subsets preserve the requested order; errors keep a named NULL entry.
selected <- dml_from_did_cells(source_fit,
  cells = data.frame(group = c(5, 9), time = c(12, 12)), dml_args = args)
benchmarks <- did_cell_apply(selected, dml_benchmark, benchmark_covariates = "z")
benchmarks$status

```

## Repeated cross-fitting

The example uses `cf.reps = 3`; omission still gives one repetition. Each cell
retains every DML repetition. The original ATT and IF remain the same, while
sensitivity moments and their joint covariance are computed separately per
partition. Any failed repetition rejects that cell and is reported by the batch.

`adapter_info$fold_id` is a unit-by-repetition matrix for multiple repetitions
(a vector for one); `rep_seeds` and `dml.att` record every repetition's seed and
auxiliary ATT. Benchmark refits preserve every partition.

```{r combine}
confidence_bounds(hybrid, cf.y = 0.05, cf.d = 0.05, combine.method = "mean")
sens_mean <- sensemakr(hybrid, cf.y = 0.05, cf.d = 0.05, combine.method = "mean")
```

The default aggregation is `"median"`. Mean/median rules combine each
repetition's endpoint and variance, including between-split variation, without
treating the splits as independent samples. The source ATT and its standard
error do not shrink as the number of repetitions increases. `sensemakr()` uses
the selected rule consistently in its tables, robustness values and default
plots. Repeated partitions do not add cross-cell aggregation or simultaneous
inference.

## Results and batching

A single fit exposes `adapter_data`, `adapter_info$cell`, source diagnostics,
and the original `external_fit`. `adapter_info$cell` records the base period,
cell/global sample sizes, original IF row indices and scaling factor.
`sensemakr()`, `dml_bounds()`, `confidence_bounds()`, `robustness_value()` and
`dml_benchmark()` accept the result.

`dml_from_did_cells()` defaults to all post-treatment source cells, including
failed source estimates. Its `models` list and `status` table retain one entry
per selected cell. Pre-treatment/reference cells appear in `excluded`.
With `on_error = "record"`, a failed cell has a named NULL entry and an error
message; warnings are retained separately. `on_error = "stop"` stops at the
first failure. Invalid source objects stop before cell fitting.

`did_cell_apply()` retains the same cell identities when applying sensitivity
or benchmark functions. A function that returns NULL successfully remains
marked `ok`. Batch confidence intervals are cellwise; they provide neither
simultaneous coverage nor an aggregated ATT sensitivity analysis.

Benchmarks remove specified design columns and refit both DRDID and DML on the
frozen sample and fold partition. They continue to work after saving and loading
the adapter with `saveRDS()`/`readRDS()`, without evaluating the original call.
Keep the intercept when removing all other regressors. Aggregating cells would
require joint cross-cell IFs and, where applicable, IFs for estimated aggregation
weights; it is deferred.
